Towards a Wide-Coverage Grammar for Swedish Using GF

### 2012-01-07 doc/report/report.tex
pres, report

\abstract{
The thesis describes the work formalising an important fragment of Swedish
in the GF formalism and to compare and evaluate the possibilities of using GF
for modelling large-scale grammars. Grammatical constructions specific to and
frequent in Swedish have been investigated and formalised in the functional
programming language.
%~~achieve a covering Swedish~~ grammar implemented
%TODO first sentence should be more, no understatements!
%in GF -- to model natural language by functional programming. 
The idea has been to combine existing language technology resources with
new techniques, and the aim to obtain a system for analysing written text.
To reach this goal, problems of computational as well as linguistic nature had to be solved.
\\We present the methods and technologies used and further inspects the possibilities
of reaching our long-term goal; to accomplish a parser for unrestricted Swedish, by
combining our rule-based grammar with statistical methods.
\\Our contribution is a wide-coverage GF lexicon, a translation of a Swedish treebank
to GF notation and a Swedish grammar implementation. The grammar is based
on the multilingual abstract syntax given in GF resource library, and now
extends it to cover constructions specific to Swedish.
We further give an example of the advantage of using dependent types when
describing grammar and syntax, in this case for dealing with reflexive pronouns.
%the linguistic phenomena

@@

\chapter{Background}  
\label{sec:background}
The work described in this thesis is part of a bigger project which aims at
using GF for parsing unrestricted Swedish. 
%TODO LREC ref?
In previous work\footnote{??}, a start was made to extend the Swedish GF grammar
and a tool for lexical acquisition was developed.
We now construct a bigger
and more expressive grammar as well as a large scale GF lexicon.
As all GF grammars, this one defines a parser, and we develop it by getting
examples, ideas and test material from the treebank Talbanken.
The project is hence heavily depending on three resources, which will be described
in this section.

@@

\end{itemize}
\subsubsection{Lexicon}
Our lexicon still lacks some parts that can be imported from Saldo.
The multiple-word verbs (\verb-vbm-) and multiple-word nouns (\verb-nm-) should
be imported to an idiom lexicon, which can be used as an extension to the main
lexicon.
For the words that we tried but failed to import, another tool for lexicon
acquisition could be used. The tool developed in the previous part of
this project\footnote{??} would be suitable.
All in all, it should be ensured that we use as much information as
possible from our resources.
%The list of not-imported can be used for adding other kinds of lexicons - idioms, have pos-tags.
%Also to analyse what's wrong, or use the words with a word guesser. The code
%has been adapted for this.
%For exapmle pronouns - dets, have combined information from Talbanken to add them in the correct
%category. How this works, look at tags. Small amount, can be done manually.

>>>>

\abstract{
The thesis describes the work of achieving a covering Swedish grammar
in the GF formalism and to compare and evaluate the possibilities of using GF
for modelling large-scale grammars. Grammatical constructions specific to and
frequent in Swedish have been investigated and formalised in the functional
programming language.
%~~achieve a covering Swedish~~ grammar implemented
%TODO first sentence should be more, no understatements!
%in GF -- to model natural language by functional programming. 
The idea has been to combine existing language technology resources with
new techniques, and the aim to obtain a system for analysing written text.
To reach this goal, problems of computational as well as linguistic nature had to be solved.
\\We present the methods and technologies used and further inspects the possibilities
of reaching our long-term goal; to accomplish a parser for unrestricted Swedish, by
combining our rule-based grammar with statistical methods.
\\Our contribution is a wide-coverage GF lexicon, a translation of a Swedish treebank
to GF notation and a Swedish grammar implementation. The grammar is based
on the multilingual abstract syntax given in GF resource library, and now
extends it to cover constructions specific to Swedish.
We further give an example of the advantage of using dependent types when
describing grammar and syntax, in this case for dealing with reflexive pronouns.
%the linguistic phenomena

@@

\chapter{Background}  
\label{sec:background}
The work described in this thesis is part of a bigger project which aims at
using GF for parsing unrestricted Swedish. 
%TODO LREC ref?
In previous work\footnote{web.student.chalmers.se/\~{}mahlberg/SwedishGrammar.pdf}, a start was made to extend the Swedish GF grammar
and a tool for lexical acquisition was developed.
We now construct a bigger
and more expressive grammar as well as a large scale GF lexicon.
As all GF grammars, this one defines a parser, and we develop it by getting
examples, ideas and test material from the treebank Talbanken.
The project is hence heavily depending on three resources, which will be described
in this section.

@@

\end{itemize}
\subsubsection{Lexicon}
Our lexicon still lacks some parts that can be imported from Saldo.
The multiple-word verbs (\verb-vbm-) and multiple-word nouns (\verb-nm-) should
be imported to an idiom lexicon, which can be used as an extension to the main
lexicon.
For the words that we tried but failed to import, another tool for lexicon
acquisition could be used. The tool developed in the previous part of
this project\footnote{web.student.chalmers.se/\~{}mahlberg/SwedishGrammar.pdf} would be suitable.
All in all, it should be ensured that we use as much information as
possible from our resources.
%The list of not-imported can be used for adding other kinds of lexicons - idioms, have pos-tags.
%Also to analyse what's wrong, or use the words with a word guesser. The code
%has been adapted for this.
%For exapmle pronouns - dets, have combined information from Talbanken to add them in the correct
%category. How this works, look at tags. Small amount, can be done manually.
